\emph{MixFishSim}: highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2018-04-06 write_up/mixfishsim_ecolmod.tex
Text updates

The fleet dynamics can be broadly categorised into three components; fleet
targeting - which determines the fleet catch efficiency and preference towards
a particular species; trip-level decisions, which determine the initial
location to be fished at the beginning of a trip; and within-trip decisions,
determining movement from one fishing spot to another within a trip.

@@

Each fleet of \textit{n} vessels is characterised by both a general efficiency,
\textit{Q}, and a population specific efficiency, ${Q_{p}}$.  Thus, the product
of these parameters affects the overall catch rates for the fleet and the
preferential targeting of one population over another. This, in combination
with the parameter choice for the step-function (as well as some randomness
from the exploratory fishing process) determines the preference of fishing
locations for the fleet.  All species prices are kept the same, across fleets,
though can be made to vary seasonally.

@@

Several studies \citep[e.g.][]{Hutton2004, Tidd2012, Girardin2015} have confirmed
past activity and past catch rates are strong predictors of fishing location
choice. For this reason, the fleet dynamics sub-model includes a learning
component, where a vessel's initial fishing location in a trip is based on
selecting from previously successful fishing locations. This is achieved by
sorting all previous fishing events in the previous trip as well as the
previous time periods in past years, and choosing randomly from the top x \% of
fishing events in value.  Simulation testing indicated that this learning
increased the mean value of catches for the vessels, over just relying on the
correlated random walk function.

@@

Fishing locations within a trip are determined by a modified random walk
process. A random walk type was chosen as it is the simplest assumption
commonly used in ecology to describe animal movement which searching for
homogeneously distributed prey about which there is uncertain knowledge. In a
random walk, movement is a stochastic process through a series of steps that
can either be equal in length or take some other functional form.  The
direction of the random walk can be correlated, a characteristic known as
`persistence', providing some overall location of directional movement
\citep{Codling2008} or uncorrelated. \\

A \textit{lévy walk} is a particular form of random walk characterised by a
heavy-tailed distribution of step-length and has received a lot of attention in
ecological theory in recent years as having shown to have very similar
characteristics as those observed by animals in nature, and being a near
optimum searching strategy for predators pursuing patchily distributed prey
\citep{Bartumeus2005, Sims2008}.  \citet{Bertrand2007} showed that Peruvian
anchovy fishermen have a stochastic search pattern similar to that observed
with a lévy walk. However, it remains a subject of debate, with the contention
that search patterns may be more simply characteristed as random walks
\citep{Sakiyama2013} with specific patterns related to the characteristics of
the prey field \citep{Sims2012}. \\

We use a modified random walk where directional change is based on a correlated
circular distribution where a favourable fishing ground is likely to be ``fished
back over" by the vessel returning in the direction it came from and step
length (i.e. the distance travelled from the current to the next fishing
location) is determined by relating recent fishing success, measured as the
summed value of fish caught, $$Rev = \sum_{s=1}^{\infty} C_{s} \cdot Pr_{s}$$
where $C_{s}$ is catch of a species, and $Pr_{s}$ price of a species, to step
distance. Here, when fishing is successful vessels remain in a similar location
and continue to exploit the local fishing grounds. When unsuccessful, they move
some distance away from the current fishing location. The movement distance
retains some degree of stochasticity, which can be controlled separately.

@@

\begin{equation*}
	StepL = e^{log(\beta_{1}) + log(\beta_{2}) - (log(\frac{\beta_{1}}{\beta_{3}}))} * Rev
\end{equation*}

@@

with $k$ the concentration parameter from the von mises distribution which we
correlate with the revenue so that $k = (Rev + 1 / RefRev) * max_{k}$,
where $max_{k}$ is the maximum concentration value, $k$, and RefRev is
parameterised as for $\beta_{3}$ in the step length function.

@@

Where several fishing vessels are exploiting the same fish population
competition is known to play an important role in local distribution of fishing
effort \citep{Gillis1998}. If several vessels are fishing on the same patch of
fish, local depletion and interference will affect fishing location choice of
the fleet as a whole \citep{Rijnsdorp2000, Poos2007a}. In order to account for
this behaviour, the fishing sub-model operates spatially on a daily time-step
so that for future days the biomass available to the fishery is reduced in the
areas fished. The cumulative effect is to make heavily fished areas less
attractive as future fishing opportunities.

@@

A fisheries-independent survey is simulated where fishing on a regular grid
begins each year at the same time for a given number of stations (a fixed
station survey design). Catches of the populations present are recorded but not
removed from the population. This provides a fishery independent snapshot of
the populations at a regular spatial distribution each year, similar to
scientific surveys undertaken by fisheries research agencies.

>>>>

The fleet dynamics can be broadly categorised into three components; fleet
targeting - which determine\replaced[id=PD]{d}{s} the fleet catch efficiency
and preference towards a particular species; trip-level decisions, which
determine\added[id=PD]{d} the initial location to be fished at the beginning of
a trip; and within-trip decisions, determining movement from one fishing spot
to another within a trip.

@@

Each fleet of \textit{n} vessels \replaced[id=PD]{was}{is} characterised by
both a general efficiency, \textit{Q}, and a population specific efficiency,
${Q_{p}}$.  Thus, the product of these parameters affect\replaced[id=PD]{s}{ed}
the overall catch rates for the fleet and the preferential targeting of one
population over another. This, in combination with the parameter choice for the
step-function \added[id=PD]{defined below} (as well as some randomness from the
exploratory fishing process) determine\replaced[id=PD]{d}{s} the preference of
fishing locations for the fleet.  All species prices
\replaced[id=PD]{were}{are} kept the same, across fleets\replaced[id=PD]{and
seasons}{, though can be made to vary seasonally}.

@@

Several studies \citep[e.g.][]{Hutton2004, Tidd2012, Girardin2015} have
confirmed past activity and past catch rates are strong predictors of fishing
location choice. For this reason, the fleet dynamics sub-model
include\replaced[id=PD]{d}{s} a learning component, where a vessel's initial
fishing location in a trip \replaced[id=PD]{wa}{i}s based on selecting from
previously successful fishing locations. This \replaced[id=PD]{wa}{i}s achieved
by sorting all previous fishing events in the previous trip as well as the
previous time periods in past years, and choosing randomly from the top 75 \%
of fishing events \replaced[id=PD]{as defined by the revenue gained}{in value}.
Simulation testing indicated that this learning increased the mean value of
catches for the vessels, over just relying on the correlated random
walk \todo{\added[id=JJ]{Correlated random walk of what}} function
\added[id=PD]{as described for the 'within trip' decisions below}.

@@

Fishing locations within a trip are determined by a modified random walk
process. A random walk type was chosen as it is the simplest assumption
commonly used in ecology to describe \added[id=PD]{optimal} animal
\deleted[id=PD]{movement which} search\replaced[id=PD]{ strategy}{ing} for
\added[id=PD]{exploiting} homogeneously distributed prey about which there is
uncertain knowledge \citep{Viswanathan1999}. In a random walk, movement is a
stochastic process through a series of steps\added[id=JJ]{. These steps have a
	length, and a direction} that can either be equal in length or take
some other functional form. The direction of the random walk can be correlated,
a characteristic known as `persistence', providing some overall
\deleted[id=PD]{location of} directional movement \citep{Codling2008}
\deleted[id=PD]{or uncorrelated}. \\

A \textit{\replaced[id=JJ]{Lévy flight} {lévy walk}} is a particular form of
random walk characterised by a heavy-tailed distribution of step-length
\replaced[id=JJ]{. The Lévy flight}{and} has received a lot of attention in
ecological theory in recent years as having shown to have very similar
characteristics as those observed by animals in nature, and being a near
optimum searching strategy for predators pursuing patchily distributed prey
\citep{Viswanathan1999, Bartumeus2005, Sims2008}.  \citet{Bertrand2007} showed
that Peruvian anchovy fishermen have a stochastic search pattern similar to
that observed with a lévy walk. However, it remains a subject of debate
\added[id=PD]{\citep[e.g. see][]{Edwards2011, Reynolds2015}}, with the
contention that search patterns may be more simply characteristed as random
walks \citep{Sakiyama2013} with specific patterns related to the
characteristics of the prey field \citep{Sims2012}. \\

We use a modified random walk where directional change is based on a correlated
circular distribution where a favourable fishing ground is likely to be
``fished back over" by the vessel returning in the direction it came from and
step length (i.e. the distance travelled from the current to the next fishing
location) is determined by \deleted[id=JJ]{relating} recent fishing success,
measured as the summed value of fish caught (revenue, $Rev$), $$Rev =
\sum_{p=1}^{P} \replaced[id=PD]{L}{C}_{p} \cdot Pr_{p}$$ where
$\replaced[id=PD]{L}{C}_{p}$ is \replaced[id=PD]{landings}{catch} of a
population $p$, and $Pr_{p}$ price of a population, to step distance. Here,
when fishing is successful vessels remain in a similar location and continue to
exploit the local fishing grounds.  When unsuccessful, they move some distance
away from the current fishing location.  The movement distance retains some
degree of stochasticity, which can be controlled separately.

@@

\begin{equation*}
	StepL = e^{log(\beta_{1}) + log(\beta_{2}) -
		(log(\frac{\beta_{1}}{\beta_{3}}))} * Rev
\end{equation*}\todo{\added[id=JJ]{So
			step length increases with increasingly gross
			revenue?}\added[id=PD]{No, the opposite}}

@@

with $k$ the concentration parameter from the von \replaced[id=JJ]{M}{m}ises
distribution which we correlate with the revenue so that $k = (Rev + 1 /
RefRev) * max_{k}$, where $max_{k}$ is the maximum concentration value, $k$,
and RefRev is parameterised as for $\beta_{3}$ in the step length function.

@@

Where several fishing vessels are exploiting the same fish population
competition is known to play an important role in local distribution of fishing
effort \citep{Gillis1998}. If several vessels are fishing on the same patch of
fish, local depletion and interference \added[id=JJ]{competition} will affect
fishing location choice of the fleet as a whole \citep{Rijnsdorp2000,
	Poos2007a}. In order to account for this behaviour, the fishing
sub-model operates spatially on a daily time-step so that for future days the
biomass available to the fishery is reduced in the areas fished. The cumulative
effect is to make heavily fished areas less attractive as future fishing
opportunities.

@@

A fisheries-independent survey is simulated where fishing on a regular grid
begins each year at the same time for a given number of stations (a fixed
station survey design). Catches of the populations \replaced[id=JJ]{at each
station}{present} are recorded but not removed from the population. This
provides a fishery independent snapshot of the populations at a regular spatial
\replaced[id=JJ]{intervals}{distribution} each year, similar to scientific
surveys undertaken by fisheries research agencies.
